---
title: "Chapter 2"
output: html_notebook
---

*Applied* section in *2.4 Exercises*

# Question 8

## (a), (b): load the data from csv file, and remove *college name* column:
```{r}
rawdata <- read.table('data/College.csv', header = TRUE, sep = ',')
head(rawdata)
college <- rawdata
rownames(college) <- college[,1]
college <- college[,-1]
head(college)
```

## (c)
i ~ iii:
```{r}
summary(college)

pairs(college[,1:10])

plot(college$Private, college$Outstate)
```
iv:
```{r}
college$Elite = "No"
college$Elite[college$Top10perc > 50] = "Yes"
college$Elite = as.factor(college$Elite)
summary(college$Elite)  # there are 27 Elite colleges here
plot(college$Elite, college$Outstate)
```
v:
```{r}
par(mfrow=c(2,2))
hist(college$Apps)
hist(college$perc.alumni, col = 2)
hist(college$S.F.Ratio, breaks = 20, col = 3)
hist(college$Expend, breaks = 100)
```

# Question 9
Make sure the *ISLR* library is installed.
Or install it with `install.packages("ISLR")`.
```{r}
library(ISLR)
data("Auto")
head(Auto)
str(Auto)
```
(a):
Quantitative predictors:
mpg, cylinder, displacement, horsepower, weight, acceleration, year.

Qualitative predictors: origin, name.

```{r}
# (b)
sapply(Auto[,1:7], range)

# (c)
sapply(Auto[,1:7], mean)
sapply(Auto[,1:7], sd)

# (d)
sbset <- Auto[-(10:85),]
sapply(sbset[,1:7], range)
sapply(sbset[,1:7], mean)
sapply(sbset[,1:7], sd)
```

```{r}
# (e)
pairs(Auto)
plot(Auto$weight, Auto$mpg)
```
The plot shows that with the increasing of the weight, the mpg is decrease.
In other words, lighter cars are more energy effective.

(f): Learn from the output graph of `pairs(Auto)` above.

# Question 10
```{r}
# (a)
library(MASS)
?Boston

# (b)
pairs(Boston)

# (c)
plot(Boston$age, Boston$crim)
```
Older house with more crime.

(d):
```{r}
library(dplyr)
head(arrange(Boston, desc(crim)))
```

(e):
```{r}
nrow(Boston[Boston$chas > 0,])
```

(f):
```{r}
median(Boston$ptratio)
```

(g):
```{r}
subset(Boston, medv == min(Boston$medv))
```

(h):
```{r}
nrow(Boston[Boston$rm > 7,])
summary(subset(Boston, rm > 8))
summary(Boston)
```

Suburbs with more average number of rooms per dwelling have lower crime rate (comparing the mean and median), etc.