---
title: "Exercise Applied in Chapter 3 of ISL"
output: html_notebook
---

8. (a).
```{r}
library(ISLR)
data(Auto)
fit <- lm(mpg ~ horsepower, data = Auto)
summary(fit)

predict(fit, data.frame(horsepower = c(98)), interval = 'confidence')
predict(fit, data.frame(horsepower = c(98)), interval = 'prediction')
```

i. There is a relationship between *mpg* and *horsepower* in data frame *Auto*, because the *p-value* is less than 0.05.

ii. The $R^2$ statistic suggests *horsepower* explains 61% of *mpg*.

iii. Negative. For the coeffecient of *horsepower* is less than 0.

iv. $mpg = 39.9 - 0.158 \times horsepower = 39.9 - 0.158 \times 98 = 24.4$

v. The associated 95% confidence interval is $[23.97, 24.96]$, the prediction interval is $[14.81, 34.12]$.


(b).
```{r}
plot(Auto$horsepower, Auto$mpg)
abline(fit)
```

(c).
```{r}
plot(predict(fit), residuals(fit))
plot(predict(fit), rstudent(fit))
par(mfrow=c(2,2))
plot(fit)
```
The patterns in the *Residuals vs Fitted* graph shows there are some non-linearity in the relationship of *horsepower* and *mpg*.

9. (a) and (b):
```{r}
plot(Auto) # equivalent to `pairs(Auto)`
cor(within(Auto, rm(name)))
```

(c).
```{r}
mfit <- lm(mpg ~ . -name, data = Auto)
summary(mfit)
```

i. Yes, there is.

ii. *weight*, *year* and *origin*.

iii. Cars become more oil-efficient as the service life continue, the *mpg* increase about 0.75 per year.

(d).
```{r}
par(mfrow=c(2,2))
plot(mfit)
par(mfrow=c(1,1))
plot(predict(mfit), rstudent(mfit))
abline(3, 0)
```
The residual plot suggests no **large** outliers. The leverage plot suggests the no. 14 observation have unusually high leverage.

In *rstudent* plot, we see there are about 4 outliers which `abs(x) > 3`, but not far from 3.

(e).
From the result of (c).ii, I choose the following 3 predictors:
```{r}
imfit1 <- lm(mpg ~ weight * year * origin, data = Auto)
summary(imfit1)
```
The result shows all the interaction items have significant effect on *mpg*.

From results of (b), we find the largest 2 correlations: cylinders vs. displacement (0.95), weight vs. displacement (0.93). So we test their interaction effects:
```{r}
imfit2 <- lm(mpg ~ cylinders * displacement + displacement * weight, data = Auto)
summary(imfit2)
```

So the interaction between displacement and weight is significant, while that between cylinders and displacement is insignificant.