---
title: "Applied Exercises of Chapter 7"
output: html_notebook
---

# Question 6

In this exercise, you will further analyze the *Wage* data set considered throughout this chapter.

## (a)

Perform polynomial regression to predict wage using age . Use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to
the results of hypothesis testing using ANOVA? Make a plot of the resulting polynomial fit to the data.

```{r}
library(ISLR)
library(boot)
set.seed(1)
degree <- 10
cv.errs <- rep(NA, degree)
for (i in 1:degree) {
  fit <- glm(wage ~ poly(age, i), data = Wage)
  cv.errs[i] <- cv.glm(Wage, fit)$delta[1]
}
```
Plot the test MSE by the degrees:
```{r}
plot(1:degree, cv.errs, xlab = 'Degree', ylab = 'Test MSE', type = 'l')
deg.min <- which.min(cv.errs)
points(deg.min, cv.errs[deg.min], col = 'red', cex = 2, pch = 19)
```

The minimum of test MSE at the degree 9. But test MSE of degree 4 is small enough.
The comparison by ANOVA (`anova(fit.1,fit.2,fit.3,fit.4,fit.5)`, on page 290, section 7.8.1) suggests degree 4 is enough.

Predict with 3 degree model:
```{r}
plot(wage ~ age, data = Wage, col = "darkgrey")
age.range <- range(Wage$age)
age.grid <- seq(from = age.range[1], to = age.range[2])
fit <- lm(wage ~ poly(age, 3), data = Wage)
preds <- predict(fit, newdata = list(age = age.grid))
lines(age.grid, preds, col = "red", lwd = 2)
```

## (b)

Fit a step function to predict wage using age , and perform cross-validation to choose the optimal number of cuts. Make a plot of the fit obtained.

```{r}
cv.errs <- rep(NA, degree)
for (i in 2:degree) {
  Wage$age.cut <- cut(Wage$age, i)
  fit <- glm(wage ~ age.cut, data = Wage)
  cv.errs[i] <- cv.glm(Wage, fit)$delta[1]
}
plot(2:degree, cv.errs[-1], xlab = 'Cuts', ylab = 'Test MSE', type = 'l')
deg.min <- which.min(cv.errs)
points(deg.min, cv.errs[deg.min], col = 'red', cex = 2, pch = 19)
```

So 8 cuts produce minimum test MSE.

Predict with 8-cuts step function:
```{r}
plot(wage ~ age, data = Wage, col = "darkgrey")
fit <- glm(wage ~ cut(age, 8), data = Wage)
preds <- predict(fit, data.frame(age = age.grid))  # both `data.frame` and `list` work
lines(age.grid, preds, col = "red", lwd = 2)
```

Understand the `cut()` function:
```{r}
res <- cut(c(1,5,2,3,8), 2)
res
length(res)
class(res[1])
```

`cut(x, k)` acts like *bin* or *binage*, turning a continuous quantitative variable into a discrete qualitative variable, by deviding the range of `x` evenly into `k` intervals.
Each interval is called a *level*.
The output of `cut(x, k)` is a vector with the same length of `x`.
Each element of output (a *factor* object) is a *level* where the corresponding input element falls in.

# Question 7

The *Wage* data set contains a number of other features not explored in this chapter, such as marital status (*maritl*), job class (*jobclass*), and others. Explore the relationships between some of these other predictors and *wage* , and use non-linear fitting techniques in order to fit flexible models to the data. Create plots of the results obtained,
and write a summary of your findings.

See introductions about regression on qulitative predictors in section 3.3.1 and 3.6.6.

Use `summary()`
```{r}
set.seed(1)
summary(Wage$maritl)
# table(Wage$maritl) the same with `summary`
summary(Wage$jobclass)
par(mfrow = c(1, 2))
plot(Wage$maritl, Wage$wage)
plot(Wage$jobclass, Wage$wage)
```

Fit wage on multiple predictors with GAM:
```{r}
library(gam)
fit1 <- gam(wage ~ lo(year, span = 0.7) + s(age, 5) + education, data = Wage)
fit2 <- gam(wage ~ lo(year, span = 0.7) + s(age, 5) + education + jobclass, data = Wage)
fit3 <- gam(wage ~ lo(year, span = 0.7) + s(age, 5) + education + maritl, data = Wage)
fit4 <- gam(wage ~ lo(year, span = 0.7) + s(age, 5) + education + jobclass + maritl, data = Wage)
anova(fit1, fit2, fit3, fit4)
```

So model *fit4* fits the best.

Plot the model:
```{r}
par(mfrow = c(2, 3))
plot(fit4, se = T, col = "blue")
```

# Question 8

Fit some of the non-linear models investigated in this chapter to the *Auto* data set. Is there evidence for non-linear relationships in this data set? Create some informative plots to justify your answer.

There's no direct way to answer if there are nonlinear realtionships between any feature in the data set. We have to choose a feature as the response, some others as predictors. And investigate if the relationship between them is linear or not.

Get the overall relationship between all pairs of the *Auto* data set:
```{r}
set.seed(1)
pairs(Auto)
```

From the plot above we can see when using *mpg* as the response, there are clear relationships between it and *cylinders*, *displacement*, *horsepower*, *weight*. Now we will find out if they are nonlinear or not.

```{r}
fit <- lm(mpg ~ poly(cylinders, 2) + poly(displacement, 5) + poly(horsepower, 5) + poly(weight, 5), data = Auto)
summary(fit)
```

The results show that there's no significant relationship between mpg and cylinders.
There's weak relationship (p-value: 0.028) between mpg and displacement.
There's strong quadratic relation between mpg and horsepower.
There's strong linear relation between mpg and weight.

Test the degree with ANOVA:
```{r}
anv1 <- gam(mpg ~ displacement + horsepower + weight, data = Auto)
anv2 <- gam(mpg ~ displacement + s(horsepower, 2) + weight, data = Auto)
anv3 <- gam(mpg ~ s(displacement, 5) + s(horsepower, 5) + s(weight, 5), data = Auto)
anova(anv1, anv2, anv3, test = 'F')
summary(anv3)
par(mfrow=c(1,3))
plot.Gam(anv3, se=TRUE, col="red")
```

According to plot of *anv3*, try quadratic with *displacement* and *horsepower*, linear with *weight*:
```{r}
anv4 <- gam(mpg ~ s(displacement, 3) + s(horsepower, 3) + weight, data = Auto)
anova(anv4, anv3, test = 'F')
par(mfrow=c(1,3))
plot(anv4, se=TRUE, col="red")
```

So model *anv4* is good enough.

Compare their test MSE:
```{r}
lm1 <- glm(mpg ~ displacement + horsepower + weight, data = Auto)
lm2 <- glm(mpg ~ poly(displacement, 3) + poly(horsepower, 3) + weight, data = Auto)
lm3 <- glm(mpg ~ poly(displacement, 5) + poly(horsepower, 5) + poly(weight, 5), data = Auto)

cv.glm(Auto, lm1, K = 10)$delta[1]
cv.glm(Auto, lm2, K = 10)$delta[1]
cv.glm(Auto, lm3, K = 10)$delta[1]
```

The results also suggest model *lm2* (same with *anv4*) is good enough.

So the conclusion of relationships with *mpg*:
mpg ~ displacement: cubic;
mpg ~ horsepower: cubic;
mpg ~ weight: linear.

# Question 9

This question uses the variables *dis* (the weighted mean of distances to five Boston employment centers) and *nox* (nitrogen oxides concentration in parts per 10 million) from the *Boston* data. We will treat *dis* as the predictor and *nox* as the response.

## (a)

Use the `poly()` function to fit a cubic polynomial regression to predict *nox* using *dis*. Report the regression output, and plot the resulting data and polynomial fits.

```{r}
library(MASS)
fit <- lm(nox ~ poly(dis, 3), data = Boston)
summary(fit)
dis.grid <- seq(min(Boston$dis), max(Boston$dis), by = 0.1)
preds <- predict(fit, list(dis = dis.grid), se = TRUE)
se.bands <- cbind(preds$fit + 2* preds$se.fit, preds$fit - 2 * preds$se.fit)
plot(nox ~ dis, data = Boston, col = "darkgrey")
lines(dis.grid, preds$fit, lwd = 2, col = "red")
matlines(dis.grid, se.bands, lwd = 1, col = "red", lty = 3)
```

Another way to plot the fit:
```{r}
preds <- predict(fit, list(dis = dis.grid))
plot(nox ~ dis, data = Boston, col = "darkgrey")
lines(dis.grid, preds, col = "red", lwd = 2)
```

Yet another way to plot the fit:
```{r}
plot(Boston$dis, Boston$nox)
points(Boston$dis, fit$fitted.values, col = 'red')
```

## (b)

Plot the polynomial fits for a range of different polynomial degrees (say, from 1 to 10), and report the associated residual sum of squares.

```{r}
rss <- rep(NA, 10)
for (i in 1:10) {
  fit <- lm(nox ~ poly(dis, i), data = Boston)
  rss[i] <- sum(fit$residuals ^ 2)
}
plot(1:10, rss, type = 'l', xlab = "Degree", ylab = "RSS")
```

## (c)

Perform cross-validation or another approach to select the optimal degree for the polynomial, and explain your results.

```{r}
testMSE <- rep(NA, 10)
for (i in 1:10) {
  fit <- glm(nox ~ poly(dis, i), data = Boston)
  testMSE[i] <- cv.glm(Boston, fit, K = 10)$delta[1]
}
plot(1:10, testMSE, type = 'l', xlab = "Degree", ylab = "Test MSE")
points(which.min(testMSE), testMSE[which.min(testMSE)], col = 'red', pch = 19)
```

So the optimal degree is 3.

## (d)

Use the `bs()` function to fit a regression spline to predict *nox* using *dis*. Report the output for the fit using four degrees of freedom. How did you choose the knots? Plot the resulting fit.

```{r}
library(splines)
dof <- 4
fit <- lm(nox ~ bs(dis, df = dof), data = Boston)
attr(bs(Boston$dis, df = dof), "knots")

preds <- predict(fit, list(dis = dis.grid), se = TRUE)
se.bands <- cbind(preds$fit + 2* preds$se.fit, preds$fit - 2 * preds$se.fit)
plot(nox ~ dis, data = Boston, col = "darkgrey")
lines(dis.grid, preds$fit, lwd = 2, col = "red")
matlines(dis.grid, se.bands, lwd = 1, col = "red", lty = 3)
```

For degree of freedom (*dof*) equals to the sum of the number of knots and degree, and the default degree is 3, there is 1 knot.

## (e)

Now fit a regression spline for a range of degrees of freedom, and plot the resulting fits and report the resulting RSS. Describe the results obtained.

```{r}
res <- c()
df.range <- 3:16
for (dof in df.range) {
  fit <- lm(nox ~ bs(dis, df = dof), data = Boston)
  res <- c(res, sum(fit$residuals ^ 2))
}
plot(df.range, res, type = 'l', xlab = 'degree of freedom', ylab = 'RSS')
```

From the results, `df = 10` is good enough.

## (f)

Perform cross-validation or another approach in order to select the best degrees of freedom for a regression spline on this data. Describe your results.

```{r}
res <- c()
for (dof in df.range) {
  fit <- glm(nox ~ bs(dis, df = dof), data = Boston)
  testMSE <- cv.glm(Boston, fit, K = 10)$delta[1]
  res <- c(res, testMSE)
}
plot(df.range, res, type = 'l', xlab = 'degree of freedom', ylab = 'Test MSE')
points(which.min(res) + 2, res[which.min(res)], col = 'red', pch = 19)
```

According to test MSE, `df = 8` is a good choice.

# Question 10

This question relates to the *College* data set.

## (a)

Split the data into a training set and a test set. Using out-of-state tuition as the response and the other variables as the predictors, perform forward stepwise selection on the training set in order
to identify a satisfactory model that uses just a subset of the predictors.

```{r}
library(ISLR)
library(leaps)
train <- sample(1: nrow(College), nrow(College)/2)
test <- -train
fit <- regsubsets(Outstate ~ ., data = College, subset = train, method = 'forward')
fit.summary <- summary(fit)
fit.summary
```

List names of 6 predictors:
```{r}
coef(fit, id = 6)
```

## (b)

Fit a GAM on the training data, using out-of-state tuition as the response and the features selected in the previous step as the predictors. Plot the results, and explain your findings.

```{r}
gam.mod <- gam(Outstate ~ Private + s(Room.Board, 5) + s(Terminal, 5) + s(perc.alumni, 5) + s(Expend, 5) + s(Grad.Rate, 5), data = College, subset = train)
par(mfrow = c(2,3))
plot(gam.mod, se = TRUE, col = 'blue')
```

Based on the shape of the fit curves, *Expend* and *Grad.Rate* are strong non-linear with the *Outstate*.

## (c)

Evaluate the model obtained on the test set, and explain the results obtained.

```{r}
preds <- predict(gam.mod, College[test, ])
RSS <- sum((College[test, ]$Outstate - preds)^2) # based on equation (3.16)
TSS <- sum((College[test, ]$Outstate - mean(College[test, ]$Outstate)) ^ 2)
1 - (RSS / TSS)   # based on equation
```

The $R^2$ statistic on test set is 0.69.

## (d)

For which variables, if any, is there evidence of a non-linear relationship with the response?

```{r}
summary(gam.mod)
```

"Anova for Nonparametric Effects" shows *Expend* has strong non-linear relationshop with the *Outstate*. *Grad.Rate* and *PhD* have moderate non-linear relationship with the *Outstate*, which coincide with the result of (b).

# Qustion 11

In Section 7.7, it was mentioned that GAMs are generally fit using a backfitting approach. The idea behind backfitting is actually quite simple. We will now explore backfitting in the context of multiple linear regression.

Suppose that we would like to perform multiple linear regression, but we do not have software to do so. Instead, we only have software to perform simple linear regression. Therefore, we take the following iterative approach: we repeatedly hold all but one coefficient estimate fixed at its current value, and update only that coefficient estimate using a simple linear regression. The process is continued until convergence—that is, until the coefficient estimates stop changing.

We now try this out on a toy example.

## (a) & (b)

Generate a response $Y$ and two predictors $X_1$ and $X_2$, with $n = 100$.
Initialize $\hat \beta_1$ to take on a value of your choice. It does not matter what value you choose.

```{r}
set.seed(1)
y <- rnorm(100)
x1 <- rnorm(100)
x2 <- rnorm(100)
beta1 <- 3.27
```

## (c)

Keeping the $\hat \beta_1$ fixed, fit the model:
$$
Y - \hat\beta_1 X_1 = \beta_0 + \beta_2 X_2 + \epsilon
$$

```{r}
a <- y - beta1 * x1
beta2 <- lm(a ~ x2)$coef[2]
```

## (d)

Keeping the $\hat \beta_2$ fixed, fit the model:
$$
Y - \hat\beta_2 X_2 = \beta_0 + \beta_1 X_1 + \epsilon
$$

```{r}
a <- y - beta2 * x2
beta1 <- lm(a ~ x1)$coef[2]
```

## (e)

Write a for loop to repeat (c) and (d) 1,000 times. Report the estimates of $\hat \beta_0$, $\hat \beta_1$, and $\hat \beta_2$ at each iteration of the for loop.
Create a plot in which each of these values is displayed, with $\hat \beta_0$, $\hat \beta_1$, and $\hat \beta_2$ each shown in a different color.

```{r}
iter <- 10
df <- data.frame(0.0, 0.27, 0.0)
names(df) <- c('beta0', 'beta1', 'beta2')
for (i in 1:iter) {
  beta1 <- df[nrow(df), 2]
  a <- y - beta1 * x1
  beta2 <- lm(a ~ x2)$coef[2]
  a <- y - beta2 * x2
  beta1 <- lm(a ~ x1)$coef[2]
  beta0 <- lm(a ~ x1)$coef[1]
  print(beta0)
  print(beta1)
  print(beta2)
  df[nrow(df) + 1,] <- list(beta0, beta1, beta2)
}
```

## (f)

Compare your answer in (e) to the results of simply performing multiple linear regression to predict $Y$ using $X_1$ and $X_2$. Use the `abline()` function to overlay those multiple linear regression coefficient estimates on the plot obtained in (e).

```{r}
plot(df$beta0, col = 'red', type = 'l')
lines(df$beta1, col = 'blue')
lines(df$beta2, col = 'green')
res <- coef(lm(y ~ x1 + x2))
abline(h = res[1], col = 'darkred', lty = 2)
abline(h = res[2], col = 'darkblue', lty = 2)
abline(h = res[3], col = 'darkgreen', lty = 2)
```

The coefficients from iterations and multiple regression are exactly the same.

## (g)

On this data set, how many backfitting iterations were required in order to obtain a “good” approximation to the multiple regression coefficient estimates?

For this data set, 3 iterations are enough to converge.

# Question 12

This problem is a continuation of the previous exercise. In a toy example with $p = 100$, show that one can approximate the multiple linear regressioncoefficient estimates by repeatedly performing simple linear regression in a backfitting procedure. How many backfitting iterations are required in order to obtain a “good” approximation to the multiple regression coefficient estimates? Create a plot to justify your answer.

Step 1: give $\hat \beta_1, \dots, \hat \beta_{99}$ arbitrary initial values.

Step 2: fit the model:
$$
Y - \hat\beta_1 X_1 - \dots - \hat\beta_{99} X_{99} = \beta_0 + \beta_{100} X_{100} + \epsilon
$$
So we have the values of $\hat \beta_0$ and $\hat \beta_{100}$.

Step 3: with this new $\hat\beta_{100}$, fit the model:
$$
Y - \hat\beta_1 X_1 - \dots - \hat\beta_{98} X_{98} - \hat\beta_{100}X_{100} = \beta_0 + \beta_{99} X_{99} + \epsilon
$$

Repeat above step until fitting the model:
$$
Y - \hat\beta_2 X_2 - \dots - \hat\beta_{100}X_{100} = \beta_0 + \beta_{1} X_{1} + \epsilon
$$

So we have new values of $\hat \beta_0$ and $\hat \beta_1$.
With this $\hat \beta_1$ fit the model in step 2 again.

Repeat this loop until $\hat\beta_0, \dots, \hat\beta_{100}$ converge.