---
title: "Lab of Chapter 10"

output:
  github_document:
    pandoc_args: --webtex
    html_preview: false
    toc: true
    toc_depth: 3

---

# Lab 1: Principal Component Analysis

Calculate mean and variance of each feature of *USArrests* data set:
```{r}
row.names(USArrests)
names(USArrests)
apply(USArrests, 2, mean)
apply(USArrests, 2, var)
```

Perform PCA on this data set with standardization beforehand:
```{r}
pr.out <- prcomp(USArrests, scale. = TRUE)
names(pr.out)
```

Note there is a dot after the `scale` in the parameter name of function `prcomp()`.

List mean and standard deviation of each feature (which are used to *standardize* a vector):
```{r}
pr.out$center
pr.out$scale
```

Note these results are exactly the same with above calculations with `apply()`, for a vector's *standard deviation* is the square root of its *variance*.

The result of `prcomp()` also has a `sdev` output (see `pr.out$sdev`), which is the standard deviation of the principal components instead of origin features.

List the *loading vectors* (each column defines a coordinate axis that the origin features have maximum variance on it):
```{r}
pr.out$rotation
```

Study the principal component score vectors:
```{r}
dim(pr.out$x)
biplot(pr.out, scale = 0)
```

Notice that this figure is a mirror image of Figure 10.1.
But their meaning keep the same: *Murder*, *Assault* and *Rape* are the (equal) dominant of PC1, while *UrbanPop* is the dominant of PC2.

The variance explained by each principal component (PC) is obtained by squaring the standard deviation of each PC. Hence we can get the *PVE* of this group PCs:
```{r}
pr.var <- pr.out$sdev ^ 2
pve <- pr.var / sum(pr.var)
pve
```

So the ï¬rst principal component explains 62.0 % of the variance in the data, the next principal component explains 24.7 % of the variance, and so forth. We can plot the PVE explained by each component, as well as the cumulative PVE, as follows:
```{r}
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained (PVE)", ylim = c(0,1), type = 'b')
plot(cumsum(pve), xlab = 'Principal Component', ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = 'b')
```

Finally a demo for function `cumsum()`:
```{r}
v <- c(1, 3, 5, 20, -5)
cumsum(v)
```

# Lab 2: Clustering

## K-Means Clustering






## Hierarchical Clustering

