---
title: "Lab of Chapter 10"

output:
  github_document:
    pandoc_args: --webtex
    html_preview: false
    toc: true
    toc_depth: 3

---

# Lab 1: Principal Component Analysis

Calculate mean and variance of each feature of *USArrests* data set:
```{r}
row.names(USArrests)
names(USArrests)
apply(USArrests, 2, mean)
apply(USArrests, 2, var)
```

Perform PCA on this data set with standardization beforehand:
```{r}
pr.out <- prcomp(USArrests, scale. = TRUE)
names(pr.out)
```

Note there is a dot after the `scale` in the parameter name of function `prcomp()`.

List mean and standard deviation of each feature (which are used to *standardize* a vector):
```{r}
pr.out$center
pr.out$scale
```

Note these results are exactly the same with above calculations with `apply()`, for a vector's *standard deviation* is the square root of its *variance*.

The result of `prcomp()` also has a `sdev` output (see `pr.out$sdev`), which is the standard deviation of the principal components instead of origin features.

List the *loading vectors* (each column defines a coordinate axis that the origin features have maximum variance on it):
```{r}
pr.out$rotation
```

Study the principal component score vectors:
```{r}
dim(pr.out$x)
biplot(pr.out, scale = 0)
```

Notice that this figure is a mirror image of Figure 10.1.
But their meaning keep the same: *Murder*, *Assault* and *Rape* are the (equal) dominant of PC1, while *UrbanPop* is the dominant of PC2.

The variance explained by each principal component (PC) is obtained by squaring the standard deviation of each PC. Hence we can get the *PVE* of this group PCs:
```{r}
pr.var <- pr.out$sdev ^ 2
pve <- pr.var / sum(pr.var)
pve
```

So the ï¬rst principal component explains 62.0 % of the variance in the data, the next principal component explains 24.7 % of the variance, and so forth. We can plot the PVE explained by each component, as well as the cumulative PVE, as follows:
```{r}
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained (PVE)", ylim = c(0,1), type = 'b')
plot(cumsum(pve), xlab = 'Principal Component', ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1), type = 'b')
```

Finally a demo for function `cumsum()`:
```{r}
v <- c(1, 3, 5, 20, -5)
cumsum(v)
```

# Lab 2: Clustering

## K-Means Clustering

Build clustering dataset:
```{r}
set.seed(2)
x <- matrix(rnorm(50 * 2), ncol = 2)
x[1:25, 1] <- x[1:25, 1] + 3
x[1:25, 2] <- x[1:25, 2] - 4
plot(x)
```

Clustering with K-Means:
```{r}
km.out <- kmeans(x, 2, nstart = 20)
km.out$cluster
```

What does the parameter *nstart* mean?
It's the random sets number in algorithm 10.1, which is studied in the following codes.

Plot the clusters:
```{r}
plot(x, col = (km.out$cluster + 1), main = 'K-Means clustering Result with K = 2', xlab = '', ylab = '', pch = 20, cex = 2)
```

Clustering with `K = 3`:
```{r}
set.seed(4)
km.out <- kmeans(x, 3, nstart = 20)
km.out
```

Compare clustering results using different *nstart* values:
```{r}
set.seed(3)
km.out <- kmeans(x, 3, nstart = 1)
km.out$tot.withinss

km.out <- kmeans(x, 3, nstart = 20)
km.out$tot.withinss
```

As *km.out$tot.withinss* is the total within-cluster sum of squares, which we seek to minimize by performing K-means clustering (Equation 10.11), the results above shows that `nstart = 20` gives better result than `nstart = 1`.

Here the `set.seed()` function assure that the initial cluster assignments in Step 1 of algorithm 10.1 can be replicated, and the K-means output will be fully reproducible.

## Hierarchical Clustering

Hierarchical clustering using complete linkage:
```{r}
hc.complete <- hclust(dist(x), method = 'complete')
hc.average <- hclust(dist(x), method = 'average')
hc.single <- hclust(dist(x), method = 'single')
par(mfrow = c(1,3))
plot(hc.complete, main = 'Complete linkage', xlab = '', ylab = '', sub = '', cex = .9)
plot(hc.average, main = 'Average linkage', xlab = '', ylab = '', sub = '', cex = .9)
plot(hc.single, main = 'Single linkage', xlab = '', ylab = '', sub = '', cex = .9)
```

Determine the cluster labels for each observation associated with a given cut of the dendrogram with the `cutree()` function:
```{r}
cutree(hc.complete, 2)
cutree(hc.average, 2)
cutree(hc.single, 2)
```

`hc.single` gave wrong results.

Cut the tree with 4:
```{r}
cutree(hc.single, 4)
```

To scale the variables before performing hierarchical clustering of the observations, we use the `scale()` function:
```{r}
xsc <- scale(x)
plot(hclust(dist(xsc), method = 'complete'), main = 'Hierarchical clustering with scaled features')
```

Clustering with correlation-based distance:
```{r}
x <- matrix(rnorm(30 * 3), ncol = 3)
dd <- as.dist(1 - cor(t(x)))
plot(hclust(dd, method = 'complete'), main = 'Complete linkage with correlation-based distance', xlab = '', sub = '')
```

# Lab 3: NCI60 Data Example

The dataset contains 64 observations, each has 6830 features:
```{r}
library(ISLR)
nci.labs <- NCI60$labs
nci.data <- NCI60$data
dim(nci.data)
nci.labs[1:4]
table(nci.labs)
```

## PCA on the NCI60 Data

Perform PCA on the dataset with scale:
```{r}
pr.out <- prcomp(nci.data, scale. = TRUE)
```

Build a color function, each color for a unique cancer type:
```{r}
cols <- function(vec) {
  colset <- rainbow(length(unique(vec)))
  return(colset[as.numeric(as.factor(vec))])
}
```

Plot the first 3 principal component score vectors(PC1 vs PC2, PC1 vs PC3):
```{r}
par(mfrow = c(1,2))
plot(pr.out$x[, 1:2], col = cols(nci.labs), pch = 19, xlab = "Z1", ylab = "Z2")
plot(pr.out$x[, c(1,3)], col = cols(nci.labs), pch = 19, xlab = "Z1", ylab = "Z3")
```

Print the summary of the PVE PCs (in the second of third line):
```{r}
summary(pr.out)
```

Plot the PVE:
```{r}
plot(pr.out)
```

Plot the PVE of each PC:
```{r}
pve <- 100 * pr.out$sdev ^ 2 / sum(pr.out$sdev ^ 2)
par(mfrow = c(1,2))
plot(pve, type = 'o', ylab = 'PVE', xlab = 'Principal Component', col = 'blue')
plot(cumsum(pve), type = 'o', ylab = 'Cumulative PVE', xlab = 'Principal Component', col = 'brown3')
```

There's an *elbow* at about the 7th PC. The first 7 PCs explained about 40% of the variances.

## Clustering the Observations of the NCI60 Data

Standardize the clustered data and perform hierarchical clustering with different linkage methods and euclidean distance as the dissimilarity measure:
```{r}
sd.data <- scale(nci.data)
par(mfrow = c(1,3))
data.dist <- dist(sd.data)
plot(hclust(data.dist), labels = nci.labs, main = 'Complete Linkage', xlab = '', ylab = '', sub = '')
plot(hclust(data.dist, method = 'average'), labels = nci.labs, main = 'Average Linkage', xlab = '', ylab = '', sub = '')
plot(hclust(data.dist, method = 'single'), labels = nci.labs, main = 'Single Linkage', xlab = '', ylab = '', sub = '')
```

Complete linkage generates more balanced dendrogram.

Use the complete linkage, and cut the tree with 4 clusters:
```{r}
hc.out <- hclust(data.dist)
hc.clusters <- cutree(hc.out, 4)
table(hc.clusters, nci.labs)
```

Plot the cut on the dendrogram that produces these four clusters:
```{r}
par(mfrow = c(1,1))
plot(hc.out, labels = nci.labs)
abline(h = 139, col = 'red')
```

Brief summary of the clustering result:
```{r}
hc.out
```

Compare K-Means and hierarchical clustering (HC) with `K=4`:
```{r}
set.seed(2)
km.out <- kmeans(sd.data, 4, nstart = 20)
table(km.out$cluster, hc.clusters)
```

Cluster no.2 of K-Means is the no.3 cluster of HC, no.3 of K-Means is no.1 of HC.
While the other 2 clusters of K-Means spread in more than 1 clusters in HC respectively.

Perform HC on the first 5 PCs instead of the full dataset:
```{r}
hc.out <- hclust(dist(pr.out$x[ , 1:5]))
plot(hc.out, labels = nci.labs, main = 'HC on First 5 Score Vectors')
table(cutree(hc.out, 4), nci.labs)
```

