---
title: "Conceptual Exercises of Chapter 9"
output: html_notebook
---

# Question 1

```{r}
x <- seq(-1, 1, by = 0.01)
y1 <- 3 * x + 1
plot(x, y1, type = 'l')
y2 <- 1 - x / 2
lines(x, y2, type = 'l', col = 'blue')
```

$1 + 3X_1 - X_2 > 0$ is the region above the black line, while $1 + 3X_1 - X_2 < 0$ is under the line.

$-2 + X_1 + 2X_2 > 0$ is the region above the blue line, while $-2 + X_1 + 2X_2 < 0$ is under the line.

# Question 2

(a) and (b):

```{r}
x <- seq(-4, 2, by = 0.01)
y1 <- 2 - sqrt(4 - (1 + x) ^ 2)
y2 <- 2 + sqrt(4 - (1 + x) ^ 2)
plot(c(x, x), c(y1, y2), type = 'l')
```

The points outside the circle denote $(1 + X_1) ^ 2 + (2 - X_2) ^ 2 \gt 4$, while The points in and on the circle denote $(1 + X_1) ^ 2 + (2 - X_2) ^ 2 \le 4$.

(c): Blue, red, blue and blue.

(d): The decision boundary can be written as:
$$
1 + 2 X_1 + X_1^2 - 4 X_2 + X_2^2 = 0
$$

which conforms the linear formula: $\beta_0 + \beta_1 X_1 + \beta_2 X_1^2 + \beta_3 X_2 + \beta_4 X_2^2 = 0$.

# Question 3

(a) and (b): Sketch the 7 observations:
```{r}
dat <- data.frame(x1 = c(3, 2, 4, 1, 2, 4, 4), x2 = c(4, 2, 4, 4, 1, 3, 1), y = c('R', 'R', 'R', 'R', 'B', 'B', 'B'))
plot(dat[1:4, 1:2], col = 'red', xlim = c(0, 4), ylim = c(0, 4))
points(dat[5:7, 1:2], col = 'blue')
x1 <- seq(0, 4, by = .01)
x2 <- x1 - 0.5
lines(x1, x2)
```

(c): Describe the classification rule for the maximal margin classifier:
Classify to Red if $1 - 2 x_1 + 2 x_2 \gt 0$, and classify to Blue otherwise.

(d): The margin is $\frac {2 - 1}{2 \sqrt 2} = \frac{\sqrt 2}4$.

(e): the support vectors are (2,2), (4,4), (2,1), (4,3).

(f): The seventh observation is far away from the hyperplane. A slight movement won't make this observation become a support vector. Hence it wouldn't affect the hyperplane.

(g): Sketch a hyperplane that is *not* the optimal separating hyperplane, and provide the equation for this hyperplane
```{r}
dat <- data.frame(x1 = c(3, 2, 4, 1, 2, 4, 4), x2 = c(4, 2, 4, 4, 1, 3, 1), y = c('R', 'R', 'R', 'R', 'B', 'B', 'B'))
plot(dat[1:4, 1:2], col = 'red', xlim = c(0, 4), ylim = c(0, 4))
points(dat[5:7, 1:2], col = 'blue')
x1 <- seq(0, 4, by = .01)
x2 <- x1 - 0.5
lines(x1, x2)
x3 <- 1.06 * x1 - 0.5
lines(x1, x3, col = 'darkgreen')
```

The equation for the new hyperplane (grreen line) is: $1 - 2.12 x_1 + 2 x_2 = 0$.

(h): Draw an additional observation on the plot so that the two classes are no longer separable by a hyperplane:
```{r}
dat <- data.frame(x1 = c(3, 2, 4, 1, 2, 4, 4, 2), x2 = c(4, 2, 4, 4, 1, 3, 1, 3), y = c('R', 'R', 'R', 'R', 'B', 'B', 'B', 'G'))
plot(dat[1:4, 1:2], col = 'red', xlim = c(0, 4), ylim = c(0, 4))
points(dat[5:7, 1:2], col = 'blue')
points(dat[8, 1:2], col = 'blue', cex = 2, pch = 20)
```

